\documentclass{article}

\usepackage[usenames,dvipsnames]{color}
\usepackage{graphicx}
\usepackage{bytefield}
\usepackage{hyperref}

\definecolor{lightgray}{gray}{0.8}
\definecolor{lightblue}{rgb}{0.92,0.92,1.0}
\definecolor{lightred}{rgb}{1.0,0.92,0.92}

\newcommand{\rcolor}{lightblue}
\newcommand{\wcolor}{lightred}
\newcommand{\reserved}{lightgray}

\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in Acrobat’s bookmarks
    pdftoolbar=true,        % show Acrobat’s toolbar?
    pdfmenubar=true,        % show Acrobat’s menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={My title},    % title
    pdfauthor={Author},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1} {key2} {key3}, % list of keywords
    pdfnewwindow=true,      % links in new window
    colorlinks=true,        % false: boxed links; true: colored links
    linkcolor=black,        % color of internal links
    citecolor=green,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}

\newcommand{\colorbitbox}[4][lrtb]{%
  \bitbox{0}{}%
  \rlap{\bitbox[#1]{#3}{\color{#2}\rule{\width}{\height}}}%
  \bitbox[#1]{#3}{#4}}
\newcommand{\colorwordbox}[4][lrtb]{%
  \bitbox{0}{}%
  \rlap{\wordbox[#1]{#3}{\color{#2}\rule{\width}{\height}}}%
  \wordbox[#1]{#3}{#4}}

\begin{document}

\begin{figure}[t]
 \centering
 \begin{bytefield}{32}
  \bitheader{0,3,4,16,19,20,24,28,32} \\
  \bitbox{16}{\nameref{field:Magic} (0x4E6F)}  &
  \bitbox{4}{\nameref{field:Version}} &
  \colorbitbox{\reserved}{4}{} &
  \bitbox{4}{\nameref{field:AddrSz}} &
  \bitbox{4}{\nameref{field:PortSz}} \\
  \wordbox{1}{\nameref{field:StatusAddr}} \\
  \wordgroupl{\rotatebox{90}{WB Cycle (Repeats)}}
   \colorbitbox{\reserved}{3}{} &
   \bitbox{1}{\rotatebox{90}{\small \nameref{field:RF}}}&
   \bitbox{12}{\nameref{field:RCount}} &
   \colorbitbox{\reserved}{3}{} &
   \bitbox{1}{\rotatebox{90}{\small \nameref{field:WF}}} &
   \bitbox{12}{\nameref{field:WCount}} \\
   \wordgroupr{\rotatebox{270}{\nameref{field:RCount} $\not= 0$}}
    \colorwordbox{\rcolor}{1}{\nameref{field:BaseRetAddr}} \\
    \colorwordbox{\rcolor}{1}{\nameref{field:ReadAddr} 1} \\
    \colorwordbox{\rcolor}{1}{\nameref{field:ReadAddr} 2} \\
    \colorwordbox[tb]{\rcolor}{1}{$\vdots$ \\[1ex]} \\
    \colorwordbox{\rcolor}{1}{\nameref{field:ReadAddr} $N$}
   \endwordgroupr \\
   \wordgroupr{\rotatebox{270}{WCount $\not= 0$}}
    \colorwordbox{\wcolor}{1}{\nameref{field:BaseWriteAddr}} \\
    \colorwordbox{\wcolor}{1}{\nameref{field:WriteVal} 1} \\
    \colorwordbox{\wcolor}{1}{\nameref{field:WriteVal} 2} \\
    \colorwordbox[tb]{\wcolor}{1}{$\vdots$ \\[1ex]} \\
    \colorwordbox{\wcolor}{1}{\nameref{field:WriteVal} $N$}
   \endwordgroupr 
  \endwordgroupl \\
 \end{bytefield}
 \caption{Etherbone Message Format}
 \label{fig:header}
\end{figure}


\section{Message Format}

An Etherbone message contains an Etherbone header followed by a 
sequence of Wishbone cycles.
Cycles are processed in the order they appear in the message.

All reserved fields should be cleared to 0.
All numeric values have bigendian format.
Packet alignment is max(32, \nameref{field:AddrSz}, \nameref{field:PortSz}).
For Etherbone version 1, this is either 32 or 64 bits.

\subsection{Etherbone Header}

The Etherbone header immediately follows the encapsulating UDP header.
If the packet alignment exceeds 32-bits, 
the header is followed by zeros padding to the alignment size.

\paragraph{Magic} \label{field:Magic}

As Etherbone is layered atop UDP, 
stray UDP messages might find their way to Etherbone endpoints.
In order to indentify the Etherbone protocol,
every valid Etherbone message includes the value 0x4E6F as a prefix.
Any packet missing this magic protocol identifier must be dropped.

\paragraph{Version} \label{field:Version}

This document describes Etherbone version 1.

Future Etherbone revisions must retain support for obsolete
Etherbone messages formats.
Etherbone responses must not use a newer message format than the request
which triggered the response.

An Etherbone endpoint which receives a request formatted to a higher version
than it supports must report failure by writing all ones to sender's
\nameref{field:StatusAddr}.

\paragraph{AddrSz} \label{field:AddrSz}

In principle, 
Wishbone buses do not have a fixed address width.
The value AddrSz specifies the number of bits used 
to represent addresses in Etherbone messages.
Only the low address bits are included,
any necessary higher bits are zero.

Supported values for AddrSz are listed below.
All other values are reserved.

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
AddrSz & Address bits \\
\hline
0 & 8 \\
1 & 16 \\
2 & 32 \\
3 & 64 \\
\hline
\end{tabular}

\paragraph{PortSz} \label{field:PortSz}
The Wishbone standard permits interconnection of slaves and masters 
with bus data widths of 8/16/32/64.
Masters must use the correct data bus width when communicating with slaves.
PortSz specifies the data bus width of the target slave.

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
PortSz & Data bus width \\
\hline
0 & 8 \\
1 & 16 \\
2 & 32 \\
3 & 64 \\
\hline
\end{tabular}

\paragraph{StatusAddr} \label{field:StatusAddr}
When StatusAddr = 0 the Etherbone message requires no response.
Otherwise, it requires one.
If the \nameref{field:AddrSz} is shorter than the packet alignment, 
the address is zero extended.

If a response is required,
the number of uncompleted phases should be written back to the originating
device at StatusAddr.
A response of 0 indicates success while
all bits high indicates an unsupported Etherbone version.

When an error occures on the Wishbone bus,
no further phases are processed.
Thus,
the number of uncompleted phases uniquely identifies the failed phase.

\subsection{Cycle Header}

Each cycle header corresponds to a single Wishbone cycle on the target bus.
If the packet alignment exceeds 32-bits, 
the header is followed by zeros padding to the alignment size.

\paragraph{RF} \label{field:RF}
If the results of Wishbone reads should be written back to a FIFO register
located at BaseRetAddr, the RF flag is set.
Otherwise read values will be written sequentially starting at the
BaseRetAddr.

\paragraph{RCount} \label{field:RCount}
This Wishbone cycle includes RCount reads.
The message will include an equal number of \nameref{field:ReadAddr} fields.

\paragraph{WF} \label{field:WF}
If the values written in this cycle should be written to a FIFO register,
the WF flag is set.
Otherwise values will be written sequentially starting at BaseWriteAddr.

\paragraph{WCount} \label{field:WCount}
This Wishbone cycle includes WCount reads.
The message will include an equal number of \nameref{field:WriteVal} fields.

\subsection{Read Data Phases}

Read data phases are processed in order.

\paragraph{BaseRetAddr} \label{field:BaseRetAddr}
This field is present only if \nameref{field:RCount} $>$ 0.
If the \nameref{field:AddrSz} is shorter than the packet alignment, 
the address is zero extended.
When it exists it indicates the address on the origin Etherbone endpoint 
to which read results should written.

\paragraph{ReadAddr} \label{field:ReadAddr}
For each ReadAddr, 
a Wishbone read data phase to that address will be executed.
If the \nameref{field:AddrSz} is shorter than the packet alignment, 
the address is zero extended.
Results are written back to \nameref{field:BaseRetAddr} on the origin endpoint.

\subsection{Write Data Phases}

Write data phases are processed in order.

\paragraph{BaseWriteAddr} \label{field:BaseWriteAddr}
This field is present only if \nameref{field:WCount} $>$ 0.
If the \nameref{field:AddrSz} is shorter than the packet alignment, 
the address is zero extended.
When it exists it indicates the address on the target Etherbone endpoint 
to which values should written.

\paragraph{WriteVal} \label{field:WriteVal}
For each WriteVal, 
a Wishbone write data phase will be executed.
If the \nameref{field:PortSz} is shorter than the packet alignment, 
the value is zero extended.
The destination address is computed from \nameref{field:BaseWriteAddr} and 
\nameref{field:WF}.

\end{document}
